---
title: "Day 05"
author: "Lizzie"
date: "2022-12-05"
knit: resrepo::knit_to_results
output:
  github_document:
    toc: yes
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Day 5: supply stacks

### Part 1

Write an algorithm to follow the instructions, moving a certain number of crates (letters) from the top of one stack (end of one vector) to the top (end) of another.

Crates are moved one at a time.

Aim: what crates end up on top of each stack?

```{r}

## read data ----
# stacks <- readLines("../data/day_05.txt", n=9)
## NB. the stacks are reversed, so the top of each stack is the start of the vectors
stacks <- list(rev(c("N","R","G","P")),
               rev(c("J","T","B","L","F","G","D","C")),
               rev(c("M","S","V")),
               rev(c("L","S","R","C","Z","P")),
               rev(c("P","S","L","V","C","W","D","Q")),
               rev(c("C","T","N","W","D","M","S")),
               rev(c("H","D","G","W","P")),
               rev(c("Z","L","P","H","S","C","M","V")),
               rev(c("R","P","F","L","W","G","Z"))
               )

## TEST DATA
# stacks <- list(c("N","Z"),
#                c("D","C","M"),
#                c("P"))
# moves <- c("move 1 from 2 to 1",
#            "move 3 from 1 to 3",
#            "move 2 from 2 to 1",
#            "move 1 from 1 to 2")

## how to index in the list
stacks[[1]][1:2]

## write and run algorithm ----
moves <- readLines("../data/day_05.txt")
moves <- moves[11:length(moves)]

stacks  # state at start

for (i in 1:length(moves)) {
  # i <- 1
  ## parse the instructions
  move <- moves[i]
  ## find the numbers
  move_n <- unlist(strsplit(move, "move | from | to "))
  move_n <- as.numeric(move_n[which(move_n != "")])
  n   <- move_n[1]  # how many to move
  sta <- move_n[2]  # from this stack
  end <- move_n[3]  # to this stack
  
  ## move j crates
  for (j in 1:n) {
    ## add the moved crate to the end stack
    # stacks[[sta]][1]  # the crate to move
    stacks[[end]] <- c(stacks[[sta]][1] , stacks[[end]])
    
    ## take away the moved crate from the start stack
    stacks[[sta]] <- stacks[[sta]][-1]
  }
  
  # cat("after step", i, "\n")
  # print(stacks)
  # cat("\n")
}

stacks  # state at end

### get answer ----

## concatenate the first letter in every stack
ans <- c()
for (i in 1:9) {
  ans <- c(ans, stacks[[i]][1])
}

paste(ans, collapse = "")

```

### Part 2

Crates are now moved as a block instead of one at a time. (*i.e. the way I wrote this code the first time round because I didn't read the question properly.*)

Aim: what crates end up on top of each stack now?

```{r}

stacks <- list(rev(c("N","R","G","P")),
               rev(c("J","T","B","L","F","G","D","C")),
               rev(c("M","S","V")),
               rev(c("L","S","R","C","Z","P")),
               rev(c("P","S","L","V","C","W","D","Q")),
               rev(c("C","T","N","W","D","M","S")),
               rev(c("H","D","G","W","P")),
               rev(c("Z","L","P","H","S","C","M","V")),
               rev(c("R","P","F","L","W","G","Z"))
               )

stacks  # state at start

for (i in 1:length(moves)) {
  # i <- 1
  ## parse the instructions
  move <- moves[i]
  ## find the numbers
  move_n <- unlist(strsplit(move, "move | from | to "))
  move_n <- as.numeric(move_n[which(move_n != "")])
  n   <- move_n[1]  # how many to move
  sta <- move_n[2]  # from this stack
  end <- move_n[3]  # to this stack
  
  ## add the moved crate to the end stack
  # stacks[[sta]][1:n]  # the crate to move
  stacks[[end]] <- c(stacks[[sta]][1:n] , stacks[[end]])
  
  ## take away the moved crate from the start stack
  stacks[[sta]] <- stacks[[sta]][-c(1:n)]
  
  # cat("after step", i, "\n")
  # print(stacks)
  # cat("\n")
}

stacks  # state at end

### get answer ----

## concatenate the first letter in every stack
ans <- c()
for (i in 1:9) {
  ans <- c(ans, stacks[[i]][1])
}

paste(ans, collapse = "")

```


End.

