---
title: "Day 02"
author: "Lizzie"
date: "2022-12-02"
knit: resrepo::knit_to_results
output:
  github_document:
    toc: yes
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Day 2: rock paper scissors

### Part 1

The first column is what your opponent is going to play: A for Rock, B for Paper, and C for Scissors. The second column must be what you play: X for Rock, Y for Paper, and Z for Scissors.

Your total score is the sum of your scores for each round. The score for a single round is the score for the shape you selected (1 for Rock, 2 for Paper, and 3 for Scissors) plus the score for the outcome of the round (0 if you lost, 3 if the round was a draw, and 6 if you won).

Aim: calculate your total score for all rounds.

```{r part-1}
library(dplyr)

## read data from file
df <- read.table("../data/day_02_1.txt")
colnames(df) <- c("opp", "me")
head(df)

## table of scores
tab <- data.frame(opp      = c("A","A","A","B","B","B","C","C","C"),
                  me       = c("X","Y","Z","X","Y","Z","X","Y","Z"),
                  my_score = c( 4 , 8 , 3 , 1 , 5 , 9 , 7 , 2 , 6 ))

## merge in
df_1 <- left_join(df, tab)

## what's my total score?
sum(df_1$my_score)

```


### Part 2

Actually, X = lose, Y = draw, Z = win.

```{r part-2}

colnames(df) <- c("opp", "my_res")

## table of scores
tab <- data.frame(opp      = c("A","A","A","B","B","B","C","C","C"),
                  me       = c("r","p","s","r","p","s","r","p","s"),
                  my_res   = c("Y","Z","X","X","Y","Z","Z","X","Y"),
                  my_score = c( 4 , 8 , 3 , 1 , 5 , 9 , 7 , 2 , 6 ))

## merge in
df_2 <- left_join(df, tab)
head(df_2)

## my total score
sum(df_2$my_score)


```


End.



